---
title: "Pilot Tests"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
```

# Experimental
<details>
  <summary>click to expand</summary>
  
## I. Soil Processing and Intitial Characterization

### sieving
sieve through 4 mm mesh.


### gravimetric moisture
```{r, eval=FALSE, echo=TRUE}
wt_tin_g
wt_tin_fmsoil_g
wt_tin_airdried_g

moisture
```

### saturation water content
```{r, eval=FALSE, echo=TRUE}
wt_setup_g # empty setup of ring+ mesh + saturation plate
wt_setup_fmsoil_g
wt_setup_fmsoil_clay_g # 5 g clay packet added
wt_setup_saturated_g # setup + saturated soil+clay system
wt_added_water_g = wt_setup_saturated_g - wt_setup_fmsoil_clay_g
```

## II. Preparing Oxalic Acid
### oxalic acid mixture
mix ?? unlabelled with ?? labelled oxalic acid
add ?? water to dissolve and make a ?? concentration solution

### adsorbing OA onto goethite
total clay needed = `r 5 * 8` g  

1. add ?? clay and ?? OA solution to each of ?? 50-mL omics-safe tubes
1. shake on vortexer for ?? minutes
1. centrifuge
1. decant supernatant
1. add ?? mL water (#1)
1. shake on vortexer for ?? minutes
1. centrifuge
1. decant supernatant
1. add ?? mL water (#2)
1. shake on vortexer for ?? minutes
1. centrifuge
1. decant supernatant
1. freeze-dry
1. gently shake to break up large sheets of dried clay
1. weigh 5 g of OA-clay into packets made of 100 um mesh.

## III. incubation setup
### requirements
- sieved soil
- prepared oxalic acid-goethite packets
- pint-size Mason jars
  - lids fitted with (a) Swagelok/compression unions, (b) rubber washers, (c) septa
- 50 mL syringe + needle
- 50 mL collection bottle + stopper + crimp seal (evacuate and seal)
- labels for jars and bottles

### procedure
**set A: wetting** (n=3)  

1. weigh 50 g ODE soil into pint-size Mason jars
1. place the oxalic acid-goethite packet in the top 1 cm of the soil.
2. add ?? mL Milli-Q water from above
3. seal the jars for 48 hours at room temperature
4. at the end of the incubation, pull 50 mL headspace and transfer to pre-evacuated bottles
    - ship gas samples to MSL
5. remove the oxalic acid-goethite packet
    - transfer to 15 mL omics-safe tube
    - freeze-dry
    - grind
    - analyze on Minerva for TC, 13C
6. mix the soil well
7. subsample for moisture
    - weigh into aluminum tin
    - dry in oven at 105 C for 24 hours
8. subsample for TC, 13C
    - weigh into 15 mL omics-safe tube
    - freeze-dry
    - grind
    - analyze on Minerva

**set B: drying** (n=3)  

1. weigh 50 g ODE soil into pint-size Mason jars
1. place the oxalic acid-goethite packet in the top 1 cm of the soil.
1. air-dry until constant weight
1. add ?? mL Milli-Q water from above
1. seal the jars and repeat steps described above



### experimental details

```{r}
expt_details = tribble(
  ~ sample, ~ empty_wt_g, ~fm_soil_added_g,
  1, NA, NA,
  2, NA, NA,
  3, NA, NA
)
```


</details>

# Results

----
<details>
  <summary>Session Info</summary>
  
Date: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>
