---
title: "Pilot Tests"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
```

# Experimental
<details>
  <summary>click to expand</summary>
  
## I. Soil Processing and Intitial Characterization

### sieving
sieve through 4 mm mesh.


### gravimetric moisture
```{r, echo=TRUE}
wt_tin_g = 0.81 
wt_tin_fmsoil_g = 9.08
wt_tin_airdried_g = 7.95

fieldmoisture = round(((wt_tin_fmsoil_g - wt_tin_airdried_g) / (wt_tin_airdried_g - wt_tin_g)) * 100, 2) 
```

field moisture = `r fieldmoisture` % 

### saturation water content
```{r, echo=TRUE}
wt_setup_g = 240.3# empty setup of ring+ mesh + saturation plate
wt_setup_fmsoil_g = 291.9
wt_setup_saturated_g = 320.7 # setup + saturated soil+clay system
wt_added_water_g = wt_setup_saturated_g - wt_setup_fmsoil_g

wt_fm_g = wt_setup_fmsoil_g - wt_setup_g
wt_od_g = wt_fm_g/((fieldmoisture/100)+1)
wt_moisture_g = wt_fm_g - wt_od_g
wt_saturation_water_g = wt_added_water_g + wt_moisture_g
```

saturation water content:  `r round((wt_saturation_water_g/wt_od_g)*10,2)` g water for 10 g ODE soil


## II. Preparing Oxalic Acid

### oxalic acid mixture
from the [isotope-calculations file](https://github.com/kaizadp/destabilization_som/blob/master/tests/isotope_enrichment_calculations.md), 

> we will mix 1.94 mg of 0.99 labelled substrate with 9.31 mg of unlabelled substrate (0.0121 enriched),
to get 11.25 mg of 0.1809 enriched substrate – for a 30-g soil core.

for 10 cores,   
 - mix  `r 1.94 * 10` mg of labelled substrate with `r 9.31 *10` mg unlabelled substrate to get `r 11.25 * 10` mg of enriched substrate.  
 - dissolve in 150 mL deionized mQ water


### adsorbing OA onto goethite
total clay needed = `r 5 * 8` g  

1. add 10 g clay + 30 mL OA solution to each 50-mL omics-safe tube (5 tubes total)
1. shake overnight at 250 rpm at room temperature
1. centrifuge for 20 minutes at 2000g
1. decant supernatant
1. add 30 mL water (#1)
1. shake on vortexer for 20 minutes
1. centrifuge for 20 minutes at 2000g
1. decant supernatant
1. add 30 mL water (#2)
1. shake on vortexer for 20 minutes
1. centrifuge for 20 minutes at 2000g
1. decant supernatant
1. freeze-dry
1. gently shake to break up large sheets of dried clay
1. weigh 5 g of OA-clay into packets made of 100 um mesh.

## III. incubation setup
### requirements
- sieved soil
- prepared oxalic acid-goethite packets
- pint-size Mason jars
  - lids fitted with (a) Swagelok/compression unions, (b) rubber washers, (c) septa
- 50 mL syringe + needle
- 50 mL collection bottle + stopper + crimp seal (evacuate and seal)
- labels for jars and bottles

### procedure
**set A: wetting** (n=3)  

1. weigh 50 g ODE soil into pint-size Mason jars
1. place the oxalic acid-goethite packet in the top 1 cm of the soil.
2. add 50 mL Milli-Q water from above
3. seal the jars for 48 hours at room temperature
4. at the end of the incubation, pull 50 mL headspace and transfer to pre-evacuated bottles
    - ship gas samples to MSL
5. remove the oxalic acid-goethite packet
    - transfer to 15 mL omics-safe tube
    - freeze-dry
    - grind
    - analyze on Minerva for TC, 13C
6. mix the soil well
7. subsample for moisture
    - weigh into aluminum tin
    - dry in oven at 105 C for 24 hours
8. subsample for TC, 13C
    - weigh into 15 mL omics-safe tube
    - freeze-dry
    - grind
    - analyze on Minerva

**set B: drying** (n=3)  

1. weigh 50 g ODE soil into pint-size Mason jars
1. place the oxalic acid-goethite packet in the top 1 cm of the soil.
1. air-dry until constant weight
1. add ?? mL Milli-Q water from above
1. seal the jars and repeat steps described above



### experimental details

```{r}
expt_details = tribble(
  ~ sample, ~ empty_wt_g, ~fm_soil_added_g,
  1, NA, NA,
  2, NA, NA,
  3, NA, NA
)
```


</details>

# Results

----
<details>
  <summary>Session Info</summary>
  
Date: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>
