---
title: "Untitled"
author: "Kaizad F. Patel"
date: "12/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HYPOTHESES

1. drying and wetting processes destabilize C via different mechanisms: desorption and cell lysis (drying), vs. increased hydrologic connectivity and improved spatial access to substrate (wetting). 
1.1. under conditions of drying, desorption of C from clay-sorbed substrate will result in [lag and release] of desorption/respiration. vs. cell lysis means we will have a pulse of more 12C vs. the newly added 13C
1. drying will have a stronger effect on soil C destabilization, compared to wetting
drying will have a higher ratio of 12/13C because the respiration will be more destabilization vs. wetting/access to native soil C. 





---

---

1. how was soil C partitioned? WEOC vs. non-weoc. as total mass, not mg/g.

1. how much of the added 13C was lost in respiration?

1. can we partition respiration C? weoc/tc vs. added C?

1. how did addition of clay influence C availability?

Addition of goethite reduced respiration

```{r}
a = 
  licor_clean %>% 
  left_join(dplyr::select(headspace_calc, core, mmol_air), by = "core") %>% 
  mutate(mmol_CO2C = CO2_ppm * mmol_air/1000000,
         umol_CO2C = round(mmol_CO2C * 1000,3)) %>% 
  dplyr::select(core, CO2_ppm, umol_CO2C) %>% 
  left_join(core_key, by = "core") 

a %>% 
  filter(type %in% c("control", "priming") &
               treatment == "1-time-zero") %>% 
  ggplot(aes(x = treatment, y = CO2_ppm, color = type))+
  geom_point(position = position_dodge(width = 0.4))

summary(
  aov(CO2_ppm ~ type, 
    data = 
      a %>% 
      filter(type %in% c("control", "priming") &
               treatment == "1-time-zero" &
               core != 40))
  )
```


## pieces needed
1. respiration: ppm, mg, delta13C
2. soil: tc, mg, delta13C
3. weoc: mg, delta13C

delta -> R -> f -> mg 13C per core
