---
title: "The role of *wetting* vs. *drying* <br> on SOM destabilization"
author: "Kaizad F. Patel"
date: "updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, ki-fonts]
    lib_dir: libs
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{css, echo = FALSE}
.remark-slide-content {
  font-size: 20px;
  padding: 20px 80px 20px 80px;
}
.remark-code, .remark-inline-code {
  background: #f0f0f0;
}
.remark-code {
  font-size: 24px;
}
.huge .remark-code { /*Change made here*/
  font-size: 200% !important;
}
.small .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

### Objectives


---

### Experiment

- 60 g field-moist soil (~58 g oven-dry equivalent) in pint size Mason jars
- labelled substrate (13C-oxalic acid, 18.66 % enriched) was added to the soil, adsorbed onto 5 g goethite
- The soil was subjected to one of the following treatments, after which the jars were sealed:
  1. wetting: 20 mL water added
  1. drying: soil allowed to air-dry until constant weight
  1. drying-rewetting: soil allowed to dry, then rewet with 20 mL water
  1. time-zero: jars sealed as soon as goethite was added
- The jars were sealed for 48 hours.

---

### Analyses

1. respiration: headspace samples were collected and analyzed for CO2 and 13C-CO2
1. WEOC (water extractable organic C): organic carbon extracted and analyzed for C content and 13C
1. total C: soil was dried and analyzed for C and 13C content

---


##### Slide 2

```{r, fig.width=5}
#readd(gg_prelim)$gg_prelim_d13c
```

---

##### Slide 2

```{r, fig.width=5}
#readd(gg_prelim)$gg_prelim_c
```

---

### Desorption Samples

.pull-left[
```{r}
readd(gg_d13c)
```
]

.pull-right[
#### Enrichment
- CO2/WEOC enrichment did not differ between wetting and drying-rewetting samples.
  - we expected d+w > wetting
  - wetting and drying-rewetting mobilized C equally? 
- soil enrichment remained constant (as expected), except for the air-dried soils
  - less C was respired when dry, so more labelled 13C accumulated in soil?
]

---

### Desorption Samples

.pull-left[
```{r}
readd(gg_c)
```
]

.pull-right[
#### total C (12C + 13C)
- more CO2 was evolved after d+w, compared to wetting
- WEOC was greatest in the air-dried samples
  - because C was not consumed, and it accumulated in the soil?

]

---

### Priming Samples

.pull-left[
```{r, fig.height=7, fig.width=3}
#readd(gg_priming)$gg_priming_d13c
```
]

.pull-right[
- respiration, WEOC, and soil were all enriched after addition of soluble 13C-oxalic acid

]

---
### Mass Balance

.pull-left[
```{r}
readd(gg_mass_balance)$gg_massbalance_desorp
```
]

.pull-right[
#### Desorption Samples
]
---
### Mass Balance

.pull-left[
```{r}
readd(gg_mass_balance)$gg_massbalance_priming
```
]


.pull-right[
#### Priming Samples
- 417 ug (`r round(417/58,2)` ug/g) 13C was added as solution
- partitioned as 0.04 ug/g in respiration and 0.08 ug/g in WEOC
]

---

### Mass Balance

.pull-left[
```{r}
readd(gg_mass_balance)$gg_mass_balance_soilc
```
]


.pull-right[
#### total C (12C + 13C)


]

---

### How did clay addition alter C?

.pull-left[
```{r, fig.width=4}
readd(gg_tzero)$gg_tzero_d13c
```
]

.pull-right[
```{r, fig.width=4}
readd(gg_tzero)$gg_tzero_c
```
]

---

Date run: `r Sys.Date()`

Slides prepared using [xaringan](https://github.com/yihui/xaringan)


.small[
```{r}
utils:::print.sessionInfo(sessionInfo()[-8])
```
]