---
title: "DOM destabilization: respiration"
---

two treatments:

1. desorption: 13C oxalic acid adsorbed to goethite, 5 g goethite+OA added to soil
2. priming: 13C oxalic acid added to soil as solution, 5 g goethite (without substrate) added to soil
3. control: 5 g goethite added to soil, but no OA, only water

Jars were sealed for 48 hours, after which headspace samples were collected and analyzed for CO2 concentration and 13C/12C composition of CO2. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      comment = "#>",
                      fig.path = ("markdown/respiration/"))

library(drake)
```

---

```{r resp_plots, fig.height=5, fig.width=6}
readd(gg_respiration)$gg_resp
readd(gg_respiration)$gg_D13C
readd(gg_respiration)$gg_R13C
readd(gg_respiration)$gg_13C_umol
```

---

### priming

How much priming was seen in the treatments?

Priming calculated as: *(CO2C_umol - 13CO2C_umol) - control_CO2C_umol*

[from Bastida et al. Nat. Comms.](https://doi.org/10.1038/s41467-019-11472-7) 


```{r priming}
readd(priming_table)
```


---

<details>
  <summary>Session Info</summary>

Date Run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>